\documentclass{llncs}

\usepackage{fullpage} % uses full 8.5x11 page

\title{Keyczar: A Cryptographic Toolkit}
\author{Arkajit Dey\inst{1} and Stephen Weis\inst{2}}

\institute{Massachusetts Institute of Technology, Cambridge, MA, USA 02139
\and
Google Inc., Mountain View, CA, USA 94043}

\begin{document}
\maketitle

\begin{abstract}
Keyczar's goal is to make it easier for application developers to safely use
cryptography. Keyczar defaults to safe algorithms, key lengths, and
modes, and prevents developers from inadvertently exposing key material. It
uses a simple, extensible key versioning system that allows developers to
easily rotate and retire keys. 
\end{abstract}

\section{Introduction and Philosophy}

The motivation for Keyczar grew out of a need to make cryptography easier to
use for developers. Developers often can make simple mistakes when using
cryptography that can create security vulnerabilities. For instance, developers
may use obsolete algorithms, weak key lengths, improper cipher modes, or
unsafely compose cryptographic operations. Another common developer mistake is
to fail to provision for key rotation or even to hard-code keys in source
code.

Keyczar's goal is to address these issues by providing a simple application
programming interface (API) for developers that handles basic cryptographic
details. Keyczar also provides a simple key versioning and management system
based on directories of human-readable flat files, which will be refered to as
\emph{keysets}. 
\section{Using KeyczarTool}\label{keytool}

All Keyczar keys are generated with the stand-alone
{\tt KeyczarTool} utility located in the {\tt com.google.keyczar} Java package.

\subsection{KeyczarTool create}
KeyczarTool must first create a new keyset using the {\tt create} command. A
newly created keyset will initially contain just a metadata file, described in
section \ref{metadata}.

{\tt KeyczarTool create} requires {\tt location} and {\tt purpose} command-line
flags that specify the location of the key set and its purpose. Valid purposes
are currently {\tt crypt} and {\tt sign}. The create command may also take
an optional {\tt name} flag to give a newly created keyset a name. If the {\tt
asymmetric} flag is specified, the newly created set will contain asymmetric
keys.

Some example {\tt create} commands:
\begin{itemize}
\item Create a symmetric signing (HMAC) keyset: \\
{\tt KeyczarTool create --location=/path/to/keyset --purpose=sign}
\item Create a symmetric crypting (AES) keyset named ``Test'': \\
{\tt KeyczarTool create --location=/path/to/keyset --purpose=crypt --name=Test}
\item Create an asymmetric signing (DSA) keyset: \\
{\tt KeyczarTool create --location=/path/to/keyset --purpose=sign --asymmetric} 
\end{itemize}

\subsection{KeyczarTool addkey}

All Keyczar keys are created using the {\tt addkey} command. This command
requites a keyset {\tt location} flag and may optionally have {\tt status},
{\tt crypter} and {\tt size} flags. Section \ref{status} describes the meaning
of the status values, but briefly they are {\it primary}, {\it active}, and {\it
scheduled for revocation}. The default status is {\it active}. User-specified
key sizes are supported, although it is recommneded that on default or larger
key  sizes are used.

The {\tt addkey} command will create a new file in the keyset directory with an
integer version number that is one greater than the currently largest version.
Version numbers start from 1 and are described in Section \ref{versions}. For
example, if the current keyset contains the key file {\tt 1}, a new key version
will be created in the file {\tt 2}. Some example {\tt addkey} commands:
\begin{itemize}
\item Create a new primary key: \\
{\tt KeyczarTool addkey --location=/path/to/keyset --status=primary}
\item Create a new active key: \\
{\tt KeyczarTool addkey --location=/path/to/keyset} 
\end{itemize}

Keyczar supports encrypted keysets. The {\tt crypter} flag may be used to
encrypt a key set when adding a new key. It will specify the location of an
existing keyset which will be used to encrypt a newly generated key: 
 
\begin{itemize}
\item Create a new active key and encrypt it with another keyset: \\
{\tt KeyczarTool addkey --location=/path/to/keyset
--crypter=/path/to/crypting/keys}
\end{itemize}

\subsection{KeyczarTool pubkey}

Public keys in Keyczar are exported from existing asymmetric key sets. The
{\tt pubkey} command requires both an existing {\tt location} flag and a {\tt
destination} where public keys will be exported. If the keyset under {\tt
location} was not created with an {\tt asymmetric} flag, then a {\tt pubkey}
command will fail. An example {\tt pubkey} command works as follows: \\
{\tt KeyczarTool pubkey --location=/path/to/keyset --destination=/path/to/dest}

\subsection{KeyczarTool promote, demote, and revoke}

The {\tt promote}, {\tt demote}, and {\tt revoke} commands are used to change
key status values. Each of these commands require a {\tt location} and {\tt
version} flag. 

Promoting an {\it active} key will raise its status to
{\it primary}, and promoting a {\it scheduled for revocation} status will make
it {\it primary}. There can only be a single {\it primary} key in a given key
set. 

Similarly, {\tt demote} will lower a {\it primary} key to {\it active},
and an {\it active} key to {\it scheduled for revocation}. The {\tt revoke}
command will only work for {\it scheduled for revocation} status values.
The {\tt revoke} command will permenantly delete key material, so should be
used with caution.

Some example {\tt promote}, {\tt demote}, and {\tt revoke} commands. Suppose
that key version 1 initially has an {\it active} status:
\begin{itemize}
\item Promote {\it active} version 1 to {\it primary}: \\
{\tt KeyczarTool promote --location=/path/to/keyset --version=1}
\item Demote {\it primary} version 1 back to {\it active}: \\
{\tt KeyczarTool demote --location=/path/to/keyset --version=1}
\item Demote {\it active} version 1 to {\it scheduled for revocation}: \\
{\tt KeyczarTool demote --location=/path/to/keyset --version=1}
\item Revoke the {\it scheduled for revocation} version 1:  
\item {\tt KeyczarTool revoke --location=/path/to/keyset --version=1}
\end{itemize}

\section{Using Java Keyczar}

The {\tt com.google.keyczar} package contains four public classes that
developers will use for cryptographic functions: {\tt Crypter}, {\tt
Encrypter}, {\tt Signer}, and {\tt Verifier}. All four classes are children
of the abstract {\tt Keyczar} class and may be constructed either with a {\tt
KeyczarReader} or a simple String location of a keyset:
\begin{itemize}
  \item Constructor: {\tt Keyczar(KeyczarReader reader)}
  \item Constructor: {\tt Keyczar(String keySetLocation)}
\end{itemize} 

\subsection{KeyczarReaders}

The {\tt KeyczarReader} interface may be implemented by any class that reads
key material from a keyset. Keyczar includes a {\tt KeyczarFileReader} class
that is used by the {\tt Keyczar(String keySetLocation)} constructor to read
files from the local disk. Developers may implement their own {\tt
KeyczarReaders} to read from arbitrary sources and pass them to {\tt
Keyczar} constructors.

\subsection{Encrypters}

The {\tt Encypter} class is only able to encrypt data. {\tt Encrypter} objects
will be initialized by passing it the location of a crypting key set.
{\tt Encrypters} can encrypt plaintext encoded in UTF-8 Strings, byte arrays, or
ByteBuffers. {\tt Encrypters} must be initialized with a keyset containing a
{\it primary} status key to encrypt any data.
\begin{itemize}
  \item {\tt String encrypt(String input)}: Encrypt a string and return a
  WebSafeBase64 output ciphertext.
  \item {\tt byte[] encrypt(byte[] input)}: Encrypt a byte array and return the
  ciphertext as a byte array.
  \item {\tt int encrypt(ByteBuffer input, ByteBuffer output)}: Encrypt the
  contents of the input buffer and write the ciphertext to the output. Return
  the number of bytes written.
\end{itemize}

\subsection{Crypters}

The {\tt Crypter} class is a child of {\tt Encrypter} and is also able to
decrypt data. {\tt Crypter} objects may only be initialized by crypting key sets
that contain keys able to decrypt, namely symmetric keys or private keys. {\tt
Crypters} cannot be initialzied with public keys. {\tt Crypters} can decrypt
ciphertext in WebSafeBase64 encoded Strings, byte arrays, or ByteBuffers.
\begin{itemize}
  \item {\tt String decrypt(String input)}: Decrypt a WebSafeBase64 string and
  return a UTF-8 encoded plaintext output.
  \item {\tt byte[] decrypt(byte[] input)}: Decrypt a ciphertext byte array and
  return the plaintext as a byte array.
  \item {\tt int decrypt(ByteBuffer input, ByteBuffer output)}: Decrypt the
  input ByteBuffer and write its output into the given output ByteBuffer.
  Return the number of bytes written.
\end{itemize}

\subsection{Verifiers}

The {\tt Verifier} class is only able to verify signatures. {\tt Verifier}
objects may be initialized by passing it the location of any signing key set.
It can verify signatures in WebSafeBase64 format, as byte arrays, or in
ByteBuffers:
\begin{itemize}
  \item {\tt boolean verify(String data, String signature)}: Verifies a
  WebSafeBase64 encoded signature on the given UTF-8 String of data. Returns a
  boolean representing whether the signature is valid.
  \item {\tt boolean verify(byte[] data, byte[] signature)}: Verifies a
  signature in a byte array on the given byte array of data. Returns a
  boolean representing whether the signature is valid. 
  \item {\tt boolean verify(ByteBuffer data, ByteBuffer signature}: Verifies a
  signauture in a ByteBuffer on the given ByteBuffer of data. Returns a boolean
  representing whether the signature is valid.
\end{itemize}

\subsection{Signers}

The {\tt Signer} class is a child of the {\tt Verifier} class and is also able
to generate signatures by signing data. {\tt Signer} objects may only be
initialized by keysets that are able to sign, namely symmetric keys and private
keys. {\tt Signers} may not be initialized by public keys and may sign data in
UTF-8 Strings, byte arrays, or ByteBuffers. {\tt Signers} must be initialized
with a keyset containing a {\it primary} status key to sign any data.
\begin{itemize}
  \item {\tt String sign(String data)}: Signs a UTF-8 string of data and
  returns a signatures as a WebSafeBase64 string.
  \item {\tt byte[] sign(byte[] data)}: Signs a byte array of data and returns
  a signature in a byte array.
  \item {\tt int sign(ByteBuffer data, ByteBuffer signature}: Signs a
  ByteBuffer of input and writes the signature to the output ByteBuffer.
  Returns the number of bytes written.
\end{itemize}

\subsection{Example: Putting KeyczarTool and Java Keyczar Together}

The following command-line commands would create a new AES encrypting key:
\begin{verbatim}
KeyczarTool create --location=/aeskeys --purpose=crypt
KeyczarTool addkey --location=/aeskeys --status=primary
\end{verbatim}

Then within a Java program, the following code would encrypt a string of data:
\begin{verbatim}
KeyczarCrypter crypter = new KeyczarCrypter(``/aeskeys'');
String ciphertext = crypter.encrypt(``Some data to encrypt'');
\end{verbatim}

This same ciphertext would be decrypted with a call to {\tt Crypter.decrypt()}:
\begin{verbatim}
String plaintext = crypter.decrypt(ciphertext);
\end{verbatim}

Using public keys is similar:
\begin{verbatim}
KeyczarTool create --location=/rsakeys --purpose=crypt --asymmetric
KeyczarTool addkey --location=/rsakeys --status=primary
KeyczarTool pubkey --location=/rsakeys --destination=/rsa-publickeys
\end{verbatim}

Then in Java, we can encrypt using the public key set, but need the private key
set to decrypt:
\begin{verbatim}
// Initialize an Encrypter with the public keys
KeyczarEncypter encrypter = new KeyczarEncrypter(``/rsa-publickeykeys'');
String ciphertext = encrypter.encrypt(``Some data to encrypt'');
... 
// Initialize a Crypter with the private keys
KeyczarCrypter crypter = new KeyczarCrypter(``/rsakeys'');
String plaintext = crypter.decrypt(ciphertext);
\end{verbatim}



Signing and verifying is similarily easy:
\begin{verbatim}
KeyczarTool create --location=/hmackeys --purpose=sign
KeyczarTool addkey --location=/hmackeys --status=primary
\end{verbatim}

In Java:
\begin{verbatim}
KeyczarSigner signer = new KeyczarSigner(``/hmackeys'');
String signature = signer.sign(``Some data to sign'');
boolean verified = signer.verify(``Some data to sign'', signature);
\end{verbatim}


\section{Using Python Keyczar}

Python Keyczar is in development.

\section{Keys}

\subsection{Statuses}\label{status}

As described in Section \ref{keytool}, Keyczar keys may have one of three
status values: {\it primary}, {\it active}, {\it scheduled for revocation}.
Keys with a primary status value are used to generate new cryptomaterial, that
is, signatures or ciphertexts. Keys of all status values may be used to verify
or decrypt legacy data, but are not used to generate new data.

The idea is that a key will start its life as a primary key and be used to
generate cryptographic output. That key can be rotated by generating a new
primary key value, and changing the existing primary key to be active. Active
keys will be kept around to verify and decrypt existing data, until they are
scheduled for revocation. A key that is scheduled for revocation is identical
to an active key, except its usage statistics are collected by default. When a
key that is scheduled for revocation is no longer used, it can be safely
revoked and removed from a key set entirely.

\subsection{Key Metadata}\label{metadata}

Each Keyczar keyset contains a metadata file that contains information about
that set's purpose, key type, and a list of key version information (see
Section \ref{versions}). Valid purposes are encrypting, verifying, encrypting
and decrypting, and signing and verifying. Having the purpose value in the
metadata prevents a key from being used for an inappropriate purpose. For
example, an RSA key that is typically used for encrypting shouldn't also be
used for signing.

These purposes correspond to their respective public and symmetric/private
keys, and are usable by the Keyczar classes specified in the table below: 

\begin{tabular}{ l | l | l }
{\bf Purpose} &  {\bf Public, Private, or Symmetric } & {\bf Class used by} \\
\hline Encrypting & Public & Encrypter or Crypter \\ \hline
Encrypting and Decrypting & Private or Symmetric & Crypter \\ \hline
Verifying & Public & Verifier or Signer \\ \hline 
Verifying and Signing & Private or Symmetric & Signer \\ \hline
\end{tabular}

Each metadata file contains a specific key type that only that set can contain.
Currently, the valid key types are:
\begin{itemize}
  \item AES
  \item HMAC-SHA1
  \item DSA Private
  \item DSA Public
  \item RSA Private
  \item RSA Public
\end{itemize}

Finally, the metadata will contain a list of version information, which is
described in Section \ref{versions}.

\subsection{Versions}\label{versions}

Each key metadata (see Section \ref{metadata}) contains a list of version
numbers corresponding to the keys contained in a given keyset. Each version
specifies an integral version number, a status of either {\it active}, {\it
primary}, or {\it scheduled for revocation}, and a boolean specifying whether
the key is exportable outside of Keyczar.

By default, key version numbers start from 1 and increase montonically.
However, key version numbers may be any positive, non-zero integral values.

\subsection{Output Headers} \label{header}

Each Keyczar output is tagged with a Keyczar-specific header. The header format
is as follows:

\begin{tabular}{| l | l |}
\hline
Format & Key Hash \\ \hline
1 byte & 4 bytes \\ \hline
\end{tabular}

\subsubsection{Version Byte}

The first byte of all output is a single byte representing the version of
keyczar that generated it. The current version byte is 1. All other values are
reserved.

\subsubsection{Key hash}\label{keyhash}

Each header includes 4-byte truncated SHA-1 hash of the raw key
material itself. The following values will be hashed for each key type:
\begin{itemize}
  \item HMAC: 4-byte integer value representing the length of the HMAC key,
  followed by the actual HMAC key bytes.
  \item AES: 4-byte integer value representing the length of the AES key,
  followed by the acutal AES key bytes, followed by the 4-byte KeyHash of the
  attached HMAC key.
  \item DSA Public: 4-byte integer value representing the length of the X.509
  representation of this key, followed by the X.509 representation.
  \item DSA Private: Same as DSA Public.
  \item RSA Public: 4-byte integer value representing the length of the X.509
  representation of this key, followed by the X.509 representation.  \item
  \item RSA Private: Same as RSA Public. 
\end{itemize}

The SHA-1 output on each of these values is truncated to the first four bytes.
For example: The truncated 4-byte truncated SHA-1 values:
\begin{enumerate}
  \item The 4-byte truncated hash of SHA-1("") = {\tt da39a3ee 5e6b4b0d 3255bfef
  95601890 afd80709} encoded in hexadecimal is {\tt da39a3ee.}
  \item The truncated hash of SHA-1("The quick brown fox jumps over the lazy
  dog") = {\tt 2fd4e1c6 7a2d28fc ed849ee1 bb76e739 1b93eb12 is 2fd4e1c6} is
  {\tt 2fd4e1c6 }.
\end{enumerate}

\subsection{Key Formats}\label{formats}

All key files are kept in JSON format. All JSON representations contain
``type'' and ``hash'' values. The hash value is the key hash described in
Section \ref{keyhash}. Each key type has its own individual fields as well.

\subsubsection{HMAC}

HMAC keys have the type ``HMAC\_SHA1" and contain an additional
``hmacKeyString'' field which is a Base64 representation of the raw HMAC key
bytes. 

\subsubsection{AES}

AES keys have the type ``AES" and contain an additional
``aesKeyString'' field which is a Base64 representation of the raw HMAC key bytes. AES keys also have an attached HMAC
key, which is kept in a ``hmacKey'' field. AES keys also have an additional
``mode'' field which currently only supports ``CBC''.

\subsubsection{Public RSA}

Public RSA keys have an ``x509'' field containing the Base64 representation of
a X.509 public key certificate. 

\subsubsection{Private RSA}

Private RSA keys contain a public RSA key, and a ``pkcs8'' field containing a
Base64 representation of a PKCS8 encoded RSA private key. 

\subsubsection{Public DSA}

Public DSA keys have an ``x509'' field containing the Base64 representation of
a X.509 public key certificate. 

\subsubsection{Private DSA}

Private DSA keys contain a public DSA key, and a ``pkcs8'' field containing a
Base64 representation of a PKCS8 encoded DSA private key. 

\section{Output Formats}

\subsection{Signatures}

A Keyczar signature consists of a header (described in Section \ref{header})
followed by a algorithm-dependent signature:

\begin{tabular}{| c | c |}
\hline
Header & Sign(header $|$ M) \\
\hline 5 bytes & Algorithm-dependent length \\ \hline
\end{tabular}

(Note: $|$ represents string concatenation.)

For example, HMAC keys will produce a 20-byte output, so have the form:

\begin{tabular}{| c | c |}
\hline
Header & HMAC-SHA1(header $|$ M) \\ \hline
5 bytes & 20 bytes \\ \hline
\end{tabular}

\subsection{Ciphertext}

A Keyczar ciphertext contains a header (described in Section \ref{header}), an
IV (if needed), the algorithm-dependent ciphertext, and a signature on the
preceding fields (for symmetric keys only):

\begin{tabular}{| c | c | c | c |}
\hline
Header & IV (if any) & Encrypt(M) & Sign(header $|$ IV $|$ M) (if any) \\ \hline
5 bytes & Algorithm-dependent & Algorithm-depndent & Algorithm-depndent \\
\hline
\end{tabular}

\vspace{1mm}

For example, 128-bit AES keys (which have an attached HMAC key) used in CBC
mode would produce the following output:

\begin{tabular}{| c | c | c | c |}
\hline
Header & IV & AES-CBC(M) & HMAC-SHA1(header $|$ IV $|$ M) \\ \hline
5 bytes & 16 bytes & Algorithm-dependent & 20 bytes \\
\hline
\end{tabular}

In contrast, RSA ciphertexts neither use IVs nor are signed. Thus, the RSA
ciphertext format is:

\begin{tabular}{| c | c | c | c |}
\hline
Header & RSA-OAEP(M)  \\ \hline
5 bytes & Single RSA block  \\
\hline
\end{tabular}


\section{Algorithm Details}

These are the default key lengths, algorithms, and padding modes used:

\begin{itemize}
  \item HMAC: Default keys are 256 bits. SHA1 used as the hash
  algorithm.
  \item AES: Default keys are 128 bits. 192 and 256 bit keys are also. CBC mode
  with PKCS\#5 padding is used by default.
  \item DSA: DSA-SHA1 signing algorithm used by default. Default key size is
  1024 bits.
  \item RSA Encryption: RSA-OAEP encryption is used. Default key size is 2048
  bits. 1024, 768, and 512 bit keys are also supported.
  \item RSA Signatures: RSA-SHA1 signing is used. Default key size is 2048
  bits. 1024, 768, and 512 bit keys are also supported.
\end{itemize}


\section{Samples and Test Vectors}

\section{Licenses and Dependencies}

Keyczar is available under an Apache 2.0 license \cite{apache2}. Java Keyczar
depends on the Google GSON package \cite{google-gson}. It also relies on the
Java's {\tt javax.crypto} package, which may not be available in  all locales
due to local laws and regulations.

Python Keyczar depends on the Python Cryptography Toolkit \cite{python-crypto}
and simplejson \cite{simplejson}. 

\section{Acknowledgements}

Thanks to Ben Laurie, Marius Schilder, and Neil Daswani for their
contributions.

\bibliography{keyczar}
\bibliographystyle{acm}

\end{document}
